Optimal strategies in mental rotation tasks

### 2013-05-13 man/final-project.tex
Update naive/von mises figure caption
Clean up bayesian quadrature section

@@

\begin{figure}[t]
  \centering
  \begin{subfigure}[b]{0.45\textwidth}
    \centering
    \includegraphics[width=\textwidth]{../figures/li_regression.pdf}
    \caption{Linear interpolation}
    \label{fig:li}
  \end{subfigure}
  \begin{subfigure}[b]{0.45\textwidth}
    \centering
    \includegraphics[width=\textwidth]{../figures/vm_regression.pdf}
    \caption{Parametric regression}
    \label{fig:vm}
  \end{subfigure}
\end{figure}

As a baseline, we defined a \textit{na\"ive} model which performs a
hill-climbing search over the similarity function until it reaches a
(possibly local) maximum. Once a maximum as been found, the model
computes an estimate of $Z$ by linearly interpolating between sampled
rotations.

@@

A more flexible strategy uses what is known as \emph{Bayesian
  quadrature} \cite{Diaconis:1988uo,OHagan:1991tx} to estimate $Z$.
Bayesian quadrature allows us to compute a posterior distrbution over
$Z$ by placing a Gaussian Process (GP) \cite{Rasmussen:2006vz} prior
on the log likelihood function $\log S$ and evaluating $\log S$ at a
particular set of points. For our purposes, we choose these points to
be a sequence of $N$ consecutive mental rotations be
$\mathbf{R}=\{R_1, \ldots{}, R_N\}$. To obtain a single estimate of
$Z$, we want compute the posterior mean:
\begin{equation}
  E[Z\ \vert \ \log S]=\int_{\log S}\left(\int_R \exp(\log{S(X_b,X_R)})p(R)\ \mathrm{d}R\right)\mathcal{N}\left(\log{S}\ \vert \ \mu_{\log S}, \Sigma_{\log S}\right)\ \mathrm{d}\log S
\end{equation}
As in \cite{Osborne:2012tm}, we can approximate this with:
\begin{equation}
  E[Z\ \vert \ \log S] \approx \int_R \mu_{S}(1 + \mu_\Delta) p(R)\ \mathrm{d}R
\end{equation} 
Where $\mu_x$ is the mean of a GP, as follows:

\begin{itemize}
\itemsep1pt\parskip0pt\parsep0pt
\item $\mu_\Delta$ is a regression over $\Delta=\mu_{\log S} - \log
  \mu_S$ given $\mathbf{R}_\Delta$, which consists of $\mathbf{R}$ and
  a set of intermediate \emph{candidate points}, as described in
  \cite{Osborne:2012tm}.
\item $\mu_S$ is the mean of a GP regression over $S$ given
  $\mathbf{R}$.
\item $\mu_{\log S}$ is the mean of a GP regression over $\log S$
  given $\mathbf{R}$.
\end{itemize}

\paragraph{Choosing the sequence of mental images}

Because we can only evaluate $S(X_a, X_R)$ at a small number $k$ of
rotations, we must make an explicit decision about which
$\mathbf{R}=\{R_1, \ldots{}, R_k\}$ to use. Ideally, we should choose
$\mathbf{R}$ such that it will give us the best approximation of
$p(X_a, X_b\vert h_1)$. One option is to choose each $R_i$ in a greedy
manner to minimize the variance of $Z$ (i.e.~give a more accurate
estimate of $S$).

>>>>

In each of these models, we denote the computed rotations to be a set
$\mathbf{R}=\{R_1, R_2, \ldots{}\}$.

@@

\begin{figure}[t]
  \centering
  \begin{subfigure}[b]{0.45\textwidth}
    \centering
    \includegraphics[width=\textwidth]{../figures/li_regression.pdf}
    \caption{Na\"ive model}
    \label{fig:li}
  \end{subfigure}
  \begin{subfigure}[b]{0.45\textwidth}
    \centering
    \includegraphics[width=\textwidth]{../figures/vm_regression.pdf}
    \caption{Parametric model}
    \label{fig:vm}
  \end{subfigure}
\end{figure}

As a baseline, we defined a \textit{na\"ive} model which performs a
hill-climbing search over the similarity function until it reaches a
(possibly local) maximum. Once a maximum as been found, the model
computes an estimate of $Z$ by linearly interpolating between sampled
rotations. Figure \ref{fig:li} shows an example of the
\textit{na\"ive} model prior to estimating $Z$.

@@

A more flexible strategy uses what is known as \emph{Bayesian
  quadrature} \cite{Diaconis:1988uo,OHagan:1991tx} to estimate $Z$.
Bayesian quadrature allows us to compute a posterior distrbution over
$Z$ by placing a Gaussian Process (GP) \cite{Rasmussen:2006vz} prior
on the function $S$ and evaluating $S$ at a particular set of
points. However, this strategy has its difficulties: while in our
case, $S$ is a non-negative likelihood function, GP regression
enforces no such constraint.

\cite{Osborne:2012tm} give a method which more often (but not
entirely) avoids this problem. Instead of placing a prior over $S$, a
prior is instead placed over $\log S$, thus ensuring that $S=e^{\log
  S}$ will be positive\footnote{In practice, as in
  \cite{Osborne:2012tm}, we use the transform of $\log(S+1)$.}:
\begin{equation*}
  E[Z\ \vert \ \log S]=\int_{\log S}\left(\int_R \exp(\log{S(X_b,X_R)})p(R)\ \mathrm{d}R\right)\mathcal{N}\left(\log{S}\ \vert \ \mu_{\log S}, \Sigma_{\log S}\right)\ \mathrm{d}\log S
\end{equation*}
where $\mu_{\log S}$ and $\Sigma_{\log S}$ are the mean and
covariance, respectively, of the GP regression over $\log S$ given
$\mathbf{R}$. We approximate this according to the method given in
\cite{Osborne:2012tm}:
\begin{equation}
  \mu_Z=E[Z\ \vert \ \log S] \approx \int_R \mu_{S}(1 + \mu_\Delta) p(R)\ \mathrm{d}R 
  \label{eq:bq-Z-mean}
\end{equation}
where $\mu_S$ is the mean of a GP regression over $S$ given
$\mathbf{R}$; and $\mu_{\Delta_c}$ is a regression over
$\Delta_c=\mu_{\log S} - \log \mu_S$ given $\mathbf{R}_c$, which
consists of $\mathbf{R}$ and a set of intermediate \emph{candidate
  points}\footnote{These candidate points do not require evaluating
  the true $S$, only the GP estimates of $S$ and $\log S$.} $c$ as
described in \cite{Osborne:2012tm}. The variance of the estimate of
$Z$ is given by:
\begin{equation}
  \sigma_Z=\mathrm{Var}(Z\ \vert \ S, \log S) = \int_R\int_{R^\prime} \mu_S(R)\mu_S(R^\prime) \Sigma_{\log S}(R, R^\prime)p(R)p(R^\prime)\ \mathrm{d}R\ \mathrm{d}R^\prime
  \label{eq:bq-Z-var}
\end{equation}

To choose $\mathbf{R}$, we pick an initial direction of rotation which
results in the higher value of $S$, and then continue rotating in that
direction until the variance in Equation \ref{eq:bq-Z-var} is low
enough that we are confident about the likelihood ratio $\ell$:
\begin{equation*}
p(\ell)\approx\frac{1}{p(X_b)}\ \mathcal{N}(Z\ \vert\ \mu_Z, \sigma_z)
\end{equation*}
Specifically, we choose $h_0$ when $p(\ell < 1)\geq 0.95$, and we
choose $h_1$ when $p(\ell > 1)\geq 0.95$. Until one of these
conditions are met (or the shape has been fully rotated), the model
will continue to compute rotations and update its estimate of $Z$.
